Automated Residual Plot Assessment with the R Package autovi and the Shiny App autovi.web

### 2025-05-09 autovi_paper.qmd
Add a section for null and boot residuals

>>>>

# Definition and simulation of null and bootstrapped residuals {#sec-null-and-boot-desc}

In the subsequent sections, we will frequently refer to null residuals and bootstrapped residuals, so it is helpful to first define and explain how they are generated.

**Null residuals** are used to generate null plots within the lineup protocol framework, serving as the foundation for the statistical testing in our automated residual plot assessment. Specifically, they represent residuals generated under the null hypothesis that the model is correctly specified. A common method for simulating null residuals in linear regression involves sampling from a normal distribution with mean zero and variance equal to the estimated variance of the error term. These simulated residuals and their corresponding plots depict what one would expect from a correctly specified model. If the true residual plot exhibits noticeable deviations from these null plots, it may suggest model misspecification.

Our computer vision model is trained to assign lower VSS to null plots and higher VSS to plots that display distinct patterns. Accordingly, statistical testing is performed by computing the proportion of null plots whose VSS equals or exceeds that of the observed residual plot. This proportion serves as a p-value for a one-sided hypothesis test.

**Bootstrapped residuals** are obtained by refitting the model on bootstrap samples, which are generated by sampling individual observations with replacement from the original dataset. The residual plots obtained from these refitted models are evaluated using the same computer vision model. The predicted VSS from the bootstrapped plots provide an empirical estimate of the variation in the VSS of the observed residual plot. By examining the proportion of bootstrapped plots that also exhibit significant violations, we can assess whether the original conclusion is robust to sampling variability.

### 2025-05-09 autovi_paper.tex
Add a section for null and boot residuals

>>>>

\section{Definition and simulation of null and bootstrapped
residuals}\label{sec-null-and-boot-desc}

In the subsequent sections, we will frequently refer to null residuals
and bootstrapped residuals, so it is helpful to first define and explain
how they are generated.

\textbf{Null residuals} are used to generate null plots within the
lineup protocol framework, serving as the foundation for the statistical
testing in our automated residual plot assessment. Specifically, they
represent residuals generated under the null hypothesis that the model
is correctly specified. A common method for simulating null residuals in
linear regression involves sampling from a normal distribution with mean
zero and variance equal to the estimated variance of the error term.
These simulated residuals and their corresponding plots depict what one
would expect from a correctly specified model. If the true residual plot
exhibits noticeable deviations from these null plots, it may suggest
model misspecification.

Our computer vision model is trained to assign lower VSS to null plots
and higher VSS to plots that display distinct patterns. Accordingly,
statistical testing is performed by computing the proportion of null
plots whose VSS equals or exceeds that of the observed residual plot.
This proportion serves as a p-value for a one-sided hypothesis test.

\textbf{Bootstrapped residuals} are obtained by refitting the model on
bootstrap samples, which are generated by sampling individual
observations with replacement from the original dataset. The residual
plots obtained from these refitted models are evaluated using the same
computer vision model. The predicted VSS from the bootstrapped plots
provide an empirical estimate of the variation in the VSS of the
observed residual plot. By examining the proportion of bootstrapped
plots that also exhibit significant violations, we can assess whether
the original conclusion is robust to sampling variability.
